---
title: "ENVS 193DS — Homework 3"
author: "Will Grimsich"
format: pdf
date: 5-27-2026
toc: TRUE
toc_depth: 2
execute:
  warning: FALSE
  message: FALSE
---
# Set Up
```{r reading in data and packages}
# reading in relevant packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)

# storing kelp/temp data in object 'kelp'
kelp <- read.csv(here("data","temp-kelp.csv"))
# storing personal data in object 'exercise'
exercise <- read.csv(here("data","exercise.csv"))
# setting global theme for plots as theme_bw
theme_set(theme_bw())
```

# Problems

## Problem 1. Giant Kelp Fronds

### a. Choosing an Appropriate Test

The appropriate test to determine the strength of the relationship between temperature and giant kelp frond elongation rate is a Pearson Correlation test. This test is appropriate because both variables of interest represent continuous, numerical data that fit into this parametric test.

### b. Creating a Visualization

```{r kelp data exploration}
ggplot(data = kelp,
       mapping = aes(x = temp_c,
                     y = kelp_elong)) +
  geom_point()
```


### c. Checking Assumptions and Running Test

### d. Communicating Results

### e. Test Implications

###f. Double Checking Work

## Problem 2. Personal Data

### a. Updating Visualizations

### b. Creating Captions

## Problem 3. Affective Visualization

### a. Describing Affective Visualization for Personal Data

### b. Creating Paper Sketch of Idea

### c. Drafting a Visualization

### d. Writing an Artist Statement

### e. Prepping materials to share in class

## Problem 4. Statistical Critique

### a. Revisiting and Summarizing

### b. Visual Clarity

### c. Aesthetic Clarity

### d. Recommendations
