---
title: "ENVS 193DS — Homework 3"
author: "Will Grimsich"
format: pdf
date: 5-27-2026
toc: TRUE
toc_depth: 2
execute:
  warning: FALSE
  message: FALSE
---
# Set Up
```{r reading in data and packages}
# reading in relevant packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(patchwork)

# storing kelp/temp data in object 'kelp'
kelp <- read.csv(here("data","temp-kelp.csv"))
# storing personal data in object 'exercise'
exercise <- read.csv(here("data","exercise.csv")) |> 
  # cleaning dataframe
  clean_names()
# setting global theme for plots as theme_bw
theme_set(theme_bw())
```

# Problems

## Problem 1. Giant Kelp Fronds

### a. Choosing an Appropriate Test

The appropriate test to determine the _strength of the relationship_ between temperature and giant kelp frond elongation rate is a Pearson Correlation test. This test is appropriate because both variables of interest represent continuous, numerical data that fit into this parametric test.

### b. Creating a Visualization

```{r kelp data exploration}
# base layer ggplot
ggplot(
  # using kelp data
  data = kelp,
  # setting axes
  mapping = aes(x = temp_c,
                y = kelp_elong)) +
  # representing individual observations as points
  geom_point(color = 'forestgreen') + 
  # relabelling axes
  labs(x = "Ambient Ocean Temperature (ºC)",
       y = "Kelp Growth Rate (cm/day)")
```

### c. Checking Assumptions and Running Test

```{r checking assumptions and running pearson correlation}
# creating temperature qq plot using base layer ggplot
qq_temp <- ggplot(
  # using kelp data
  data = kelp,
  # setting sample to temperature
  mapping = aes(sample = temp_c)) +
  # adding QQ line
  geom_qq_line() +
  # adding qq points and coloring blue
  geom_qq(col = 'blue') +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Ocean Temperature")
# creating kelp qq plot using base layer ggplot
qq_kelp <- ggplot(
  # using kelp data
  data = kelp,
  # setting sample to temperature
  mapping = aes(sample = kelp_elong)) +
  # adding QQ line
  geom_qq_line() +
  # adding qq points and coloring blue
  geom_qq(col = 'forestgreen') +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Kelp Growth Rate")
# creating temperature histogram using base layer ggplot
histogram_temp <- ggplot(
  # using data from object 'kelp'
  data = kelp,
  # setting axis as temperature
  mapping = aes(x = temp_c)) +
  # creating blue histogram
  geom_histogram(fill = 'blue',
                 # using rice rule for bin number
                 bins = 6) +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Ocean Temperature")

# creating kelp histogram using base layer ggplot
histogram_kelp <- ggplot(
  # using data from object 'kelp'
  data = kelp,
  # setting axis as kelp elongation rate
  mapping = aes(x = kelp_elong)) +
  # creating green histogram
  geom_histogram(fill = 'forestgreen',
                 # using rice rule for bin number
                 bins = 6) +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Kelp Growth Rate")

# displaying above plots in 2x2 grid using patchwork package
(qq_temp + qq_kelp)/ (histogram_temp + (histogram_kelp))

# displaying 10 random rows from kelp data
slice_sample(kelp, n = 10)

# performing a correlation test between kelp growth rate and ocean temperature
cor.test(
  # formula: variable, variable
  kelp$kelp_elong, kelp$temp_c,
  # specifying pearson's correlation
  method = "pearson")
```

I checked the following assumptions necessary for a pearson correlation:

1. Continuity
2. Normal distribution
3. Linear Relationship
4. Independent Observations

I checked these assumptions using the following methods:

1. I displayed 10 random rows and checked for decimal measurements. I also thought critically about what kind of data I'm using; Both temperature and growth rates are infinitely divisible.
2. I visually assessed the normality of the distribution of both variables using QQ plots and histograms.
3. I visually assessed linearity between both variables using a scatterplot.
4. I carefully read through the metadata where it explicitly states "giant kelp _individuals_".

Through these methods I assessed the data to be both continuous and relatively normally distributed. While the temperature data are heavy-tailed, it is still unimodal and the majority of QQ points fall along the reference line. A negative, linear relationship between both variables was evident using a scatterplot. I am trusting that the observations are independent of eachother since I am told each individual is tracked, and no repeated observations are mentioned.

### d. Communicating Results

To evaluate the strength of the relationship between ocean temperature and giant kelp frond growth rate, I utilized a pearson's product-moment correlation test.

We found a moderate, negative relationship between ocean temperature and giant kelp frond growth rate (Pearson's r = -0.688, t(30) = -5.2, p < 0.001, $\alpha$ = 0.05). Therefore, we reject the null hypothesis that there is no relationship between ocean temperature and kelp growth rates.

### e. Test Implications

Through a Pearson's correlation test, I found there to be a moderately strong correlation between ocean temperature and giant kelp frond growth rates. Growth rates displayed a negative linear relationship with ocean temperature, with kelp thriving in temperatures below 16ºC. The lower boundary of my test was 14ºC, therefore further research in colder environments may be warranted.

###f. Double Checking Work

```{r spearman correlation}
# performing a correlation test between kelp growth rate and ocean temperature
cor.test(
  # formula: variable, variable
  kelp$temp_c, kelp$kelp_elong,
  # specifying spearman's rank correlation test
         method = "spearman")
```

Both Spearman and Pearson correlation tests would lead me to reject the null hypothesis since both tests have p-values much less than my significance level of 0.05 (p $\approx$ 1.3e-05). Both tests also describe the relationship between ocean temperature and giant kelp frond growth rate as negative, linear and moderately influential (Pearson's r = -0.688, Spearman's $\rho$ = -0.689).

## Problem 2. Personal Data

### a. Updating Visualizations

```{r exercise vs screen time scatterplot}
ggplot(
  # using exercise dataframe
  data = exercise,
  aes(
    # continuous Predictor Variable
    x = exercise_time, 
    # response Variable
    y = screen_time) 
) +
  # point layer representing individual observations
  geom_point(color = 'navyblue',
             # increasing size for aesthetics
             size = 2,
             shape = 21) + 
  # labelling axes
  labs(x = "Time Spent Exercising", 
       y = "Screen Time (minutes)",
       subtitle = "Most Recent Observation: 5-26-2026") +
  # using a non-default theme
  theme_bw() + 
  # removing legend
  theme(legend.position = "none") 

cor.test(exercise$exercise_time, exercise$screen_time,
         method = "pearson")
```



```{r}
ggplot(
  # using exercise dataframe
  data = exercise, 
  # mapping variables to coordinates and color
  mapping = aes(
    # categorical predictor variable on x-axis
    x = exercise_type, 
    # response variable on y-axis
    y = screen_time, 
    # mapping point color to exercise groups
    color = exercise_type)
) + 
  # jitter layer to spread out overlapping observations
  geom_jitter(
    # restricting vertical movement to show exact data values
    height = 0, 
    # slightly shifting points horizontally to prevent stacking
    width = 0.1, 
    # setting point type to a filled circle border
    shape = 21, 
  ) + 
  # manually assigning colors to categories
  scale_color_manual(
    values = c("Gym" = "red", "Gym + Run" = 'purple', "Run" = 'blue')
  ) + 
  # removing legend
  theme(legend.position = "none") + 
  # labelling axes
  labs(x = "Exercise Type",
       y = "Screen Time (minutes)",
       subtitle = "Most Recent Observation: 5-26-2026")
```


### b. Creating Captions

## Problem 3. Affective Visualization

### a. Describing Affective Visualization for Personal Data

### b. Creating Paper Sketch of Idea

### c. Drafting a Visualization

### d. Writing an Artist Statement

### e. Prepping materials to share in class

## Problem 4. Statistical Critique

### a. Revisiting and Summarizing

### b. Visual Clarity

### c. Aesthetic Clarity

### d. Recommendations
