---
title: "ENVS 193DS — Homework 3"
author: "Will Grimsich"
format: pdf
date: 5-27-2026
toc: TRUE
toc_depth: 2
execute:
  warning: FALSE
  message: FALSE
---
# Set Up
```{r reading in data and packages}
# reading in relevant packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(patchwork)

# storing kelp/temp data in object 'kelp'
kelp <- read.csv(here("data","temp-kelp.csv"))
# storing personal data in object 'exercise'
exercise <- read.csv(here("data","exercise.csv"))
# setting global theme for plots as theme_bw
theme_set(theme_bw())
```

# Problems

## Problem 1. Giant Kelp Fronds

### a. Choosing an Appropriate Test

The appropriate test to determine the strength of the relationship between temperature and giant kelp frond elongation rate is a Pearson Correlation test. This test is appropriate because both variables of interest represent continuous, numerical data that fit into this parametric test.

### b. Creating a Visualization

```{r kelp data exploration}
# base layer ggplot
ggplot(
  # using kelp data
  data = kelp,
  # setting axes
  mapping = aes(x = temp_c,
                y = kelp_elong)) +
  # representing individual observations as points
  geom_point(color = 'forestgreen') + 
  # relabelling axes
  labs(x = "Ambient Ocean Temperature (ºC)",
       y = "Kelp Growth Rate (cm/day)")
```

### c. Checking Assumptions and Running Test

```{r}
# creating temperature qq plot using base layer ggplot
qq_temp <- ggplot(
  # using kelp data
  data = kelp,
  # setting sample to temperature
  mapping = aes(sample = temp_c)) +
  # adding QQ line
  geom_qq_line() +
  # adding qq points and coloring blue
  geom_qq(col = 'blue') +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Ocean Temperature")

# creating kelp qq plot using base layer ggplot
qq_kelp <- ggplot(
  # using kelp data
  data = kelp,
  # setting sample to temperature
  mapping = aes(sample = kelp_elong)) +
  # adding QQ line
  geom_qq_line() +
  # adding qq points and coloring blue
  geom_qq(col = 'forestgreen') +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Kelp Growth Rate")

# creating temperature histogram using base layer ggplot
histogram_temp <- ggplot(
  # using data from object 'kelp'
  data = kelp,
  # setting axis as temperature
  mapping = aes(x = temp_c)) +
  # creating blue histogram and using rice rule for bin number
  geom_histogram(fill = 'blue',
                 bins = 6) +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Ocean Temperature")

# creating kelp histogram using base layer ggplot
histogram_kelp <- ggplot(
  # using data from object 'kelp'
  data = kelp,
  # setting axis as kelp elongation rate
  mapping = aes(x = kelp_elong)) +
   # creating green histogram and using rice rule for bin number
  geom_histogram(fill = 'forestgreen',
                 bins = 6) +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Kelp Growth Rate")

# displaying above plots in 2x2 grid using patchwork package
(qq_temp + qq_kelp)/ (histogram_temp + (histogram_kelp))

# displaying 10 random rows from data
slice_sample(kelp, n = 10)
```

### d. Communicating Results



### e. Test Implications



###f. Double Checking Work



## Problem 2. Personal Data

### a. Updating Visualizations

### b. Creating Captions

## Problem 3. Affective Visualization

### a. Describing Affective Visualization for Personal Data

### b. Creating Paper Sketch of Idea

### c. Drafting a Visualization

### d. Writing an Artist Statement

### e. Prepping materials to share in class

## Problem 4. Statistical Critique

### a. Revisiting and Summarizing

### b. Visual Clarity

### c. Aesthetic Clarity

### d. Recommendations
