---
title: "ENVS 193DS — Homework 3"
author: "Will Grimsich"
format: pdf
date: 5-27-2026
toc: TRUE
toc_depth: 2
execute:
  warning: FALSE
  message: FALSE
---

# Set Up

```{r reading in data and packages}
# reading in relevant packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(patchwork)

# storing kelp/temp data in object 'kelp'
kelp <- read.csv(here("data","temp-kelp.csv"))
# storing personal data in object 'exercise'
exercise <- read.csv(here("data","exercise.csv")) |> 
  # cleaning dataframe
  clean_names()
# setting global theme for plots as theme_bw
theme_set(theme_bw())
```

# Problems

## Problem 1. Giant Kelp Fronds

### a. Choosing an Appropriate Test

The appropriate test to determine the *strength of the relationship* between temperature and giant kelp frond elongation rate is a Pearson Correlation test. This test is appropriate because both variables of interest represent continuous, numerical data that fit into this parametric test.

### b. Creating a Visualization

```{r kelp data exploration}
# base layer ggplot
ggplot(
  # using kelp data
  data = kelp,
  # setting axes
  mapping = aes(x = temp_c,
                y = kelp_elong)) +
  # representing individual observations as points
  geom_point(color = 'forestgreen') + 
  # relabelling axes
  labs(x = "Ambient Ocean Temperature (ºC)",
       y = "Kelp Growth Rate (cm/day)")
```

### c. Checking Assumptions and Running Test

```{r checking assumptions and running pearson correlation}
# creating temperature qq plot using base layer ggplot
qq_temp <- ggplot(
  # using kelp data
  data = kelp,
  # setting sample to temperature
  mapping = aes(sample = temp_c)) +
  # adding QQ line
  geom_qq_line() +
  # adding qq points and coloring blue
  geom_qq(col = 'blue') +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Ocean Temperature")
# creating kelp qq plot using base layer ggplot
qq_kelp <- ggplot(
  # using kelp data
  data = kelp,
  # setting sample to temperature
  mapping = aes(sample = kelp_elong)) +
  # adding QQ line
  geom_qq_line() +
  # adding qq points and coloring blue
  geom_qq(col = 'forestgreen') +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Kelp Growth Rate")
# creating temperature histogram using base layer ggplot
histogram_temp <- ggplot(
  # using data from object 'kelp'
  data = kelp,
  # setting axis as temperature
  mapping = aes(x = temp_c)) +
  # creating blue histogram
  geom_histogram(fill = 'blue',
                 # using rice rule for bin number
                 bins = 6) +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Ocean Temperature")

# creating kelp histogram using base layer ggplot
histogram_kelp <- ggplot(
  # using data from object 'kelp'
  data = kelp,
  # setting axis as kelp elongation rate
  mapping = aes(x = kelp_elong)) +
  # creating green histogram
  geom_histogram(fill = 'forestgreen',
                 # using rice rule for bin number
                 bins = 6) +
  # removing axis labels and adding title
  labs(x = "",
       y = "",
       title = "Kelp Growth Rate")

# displaying above plots in 2x2 grid using patchwork package
(qq_temp + qq_kelp)/ (histogram_temp + (histogram_kelp))

# displaying 10 random rows from kelp data
slice_sample(kelp, n = 10)

# performing a correlation test between kelp growth rate and ocean temperature
cor.test(
  # formula: variable, variable
  kelp$kelp_elong, kelp$temp_c,
  # specifying pearson's correlation
  method = "pearson")
```

I checked the following assumptions necessary for a pearson correlation:

1.  Continuity
2.  Normal distribution
3.  Linear Relationship
4.  Independent Observations

I checked these assumptions using the following methods:

1.  I displayed 10 random rows and checked for decimal measurements. I also thought critically about what kind of data I'm using; Both temperature and growth rates are infinitely divisible.
2.  I visually assessed the normality of the distribution of both variables using QQ plots and histograms.
3.  I visually assessed linearity between both variables using a scatterplot.
4.  I carefully read through the metadata where it explicitly states "giant kelp *individuals*".

Through these methods I assessed the data to be both continuous and relatively normally distributed. While the temperature data are heavy-tailed, it is still unimodal and the majority of QQ points fall along the reference line. A negative, linear relationship between both variables was evident using a scatterplot. I am trusting that the observations are independent of eachother since I am told each individual is tracked, and no repeated observations are mentioned.

### d. Communicating Results

To evaluate the strength of the relationship between ocean temperature and giant kelp frond growth rate, I utilized a pearson's product-moment correlation test.

We found a moderate, negative relationship between ocean temperature and giant kelp frond growth rate (Pearson's r = -0.688, t(30) = -5.2, p \< 0.001, $\alpha$ = 0.05). Therefore, we reject the null hypothesis that there is no relationship between ocean temperature and kelp growth rates.

### e. Test Implications

Through a Pearson's correlation test, I found there to be a moderately strong correlation between ocean temperature and giant kelp frond growth rates. Growth rates displayed a negative linear relationship with ocean temperature, with kelp thriving in temperatures below 16ºC. The lower boundary of my test was 14ºC, therefore further research in colder environments may be warranted.

###f. Double Checking Work

```{r spearman correlation}
# performing a correlation test between kelp growth rate and ocean temperature
cor.test(
  # formula: variable, variable
  kelp$temp_c, kelp$kelp_elong,
  # specifying spearman's rank correlation test
         method = "spearman")
```

Both Spearman and Pearson correlation tests would lead me to reject the null hypothesis since both tests have p-values much less than my significance level of 0.05 (p $\approx$ 1.3e-05). Both tests also describe the relationship between ocean temperature and giant kelp frond growth rate as negative, linear and moderately influential (Pearson's r = -0.688, Spearman's $\rho$ = -0.689).

## Problem 2. Personal Data

### a. Updating Visualizations

```{r exercise vs screen time scatterplot}
# base layer ggplot
ggplot(
  # using exercise dataframe
  data = exercise,
  aes(
    # continuous Predictor Variable
    x = exercise_time, 
    # response Variable
    y = screen_time) 
) +
  # point layer representing individual observations
  geom_point(color = 'navyblue',
             # increasing size for aesthetics
             size = 2,
             # changing shape for visual clarity
             shape = 21) + 
  # labelling axes, adding titles
  labs(x = "Time Spent Exercising (minutes)", 
       y = "Screen Time (minutes)",
       title = "Exercise vs Screen Time",
       subtitle = "Most Recent Observation: 5-26-2026") +
  # removing legend and gridlines
  theme(legend.position = "none",
        panel.grid = element_blank()) 
```

```{r}
# base layer ggplot
ggplot(
  # using exercise dataframe
  data = exercise, 
  # mapping variables to coordinates and color
  mapping = aes(
    # categorical predictor variable on x-axis
    x = exercise_type, 
    # response variable on y-axis
    y = screen_time, 
    # mapping point color to exercise groups
    color = exercise_type)
) + 
  # jitter layer to spread out overlapping observations
  geom_jitter(
    # restricting vertical movement to show exact data values
    height = 0, 
    # slightly shifting points horizontally to prevent stacking
    width = 0.1, 
    # setting point type to a filled circle border
    shape = 21, 
  ) + 
  # manually assigning colors to categories, including NA
  scale_color_manual(
    values = c("Gym" = "red",
               "Gym + Run" = 'purple',
               "Run" = 'blue'),
    na.value = 'darkgreen'
  ) + 
  # removing legend and gridlines
  theme(legend.position = "none",
        panel.grid = element_blank()) + 
  # labelling axes, adding titles
  labs(x = "Exercise Type",
       y = "Screen Time (minutes)",
       title = "Screen Time by Exercise Type",
       subtitle = "Most Recent Observation: 5-26-2026") +
  # adding mean and 95% confidence interval point range
  stat_summary(geom = "pointrange",
               fun.data = "mean_se",
               # dodging jitterplot
               position = position_nudge(x = 0.25),
               # coloring black for visual clarity
               color = 'black')
```

### b. Creating Captions

**Fig. 1** Daily screen time displays a random relationship to time spent exercising. Blue circles represent individual observations of screen time as they relate to time spent exercising.

**Fig. 2** Daily screen time varies by type of exercise that day. Circles represent individual observations of screen time, categorized by the type of exercise that day. Black points represent average screen time values for that category. Black lines represent 95% confidence intervals pertaining to averages.

## Problem 3. Affective Visualization

### a. Describing Affective Visualization for Personal Data

One idea for an affective visualization I have for my data is to order the outputs in a way that displays the contrast in screen time between days where I do exercise and days where I don't get exercise. This could be done by layering horizontal lines representing average screen time, grouping exercise days together and leaving the non-exercise data as an outlying group. Furthermore, to represent the clutter and randomness of my data, I could incorporate a creative element such as unrealistic midpoints of the lines along the graph.

### b. Creating Paper Sketch of Idea

![Sketch](Exercise_Type.pdf){width="50%" out-width="50%" fig-align="center" angle="90"}

### c. Drafting a Visualization

![Rough Draft](Exercise_Draft.pdf)

### d. Writing an Artist Statement

- I am displaying the differences in my average screen time observed as they relate to different forms of exercise.

- I took influence from the Environmental Graphiti gallery, which displayed street-style art with meaningful data behind the scenes, still visually depicted.

- My work will either be a digital drawing or perhaps a watercolor/ painting.

- My rough draft was created in goodnotes where I drew out my sketch onto a digital medium, added some color and then exported as a PDF

### e. Prepping materials to share in class

[Link](https://docs.google.com/presentation/d/11-iDbN5yY4rNqEIdqS34uUQAD6B1dgElv4Noqz4Cn5I/edit?usp=sharing) to slide presentation.

## Problem 4. Statistical Critique

### a. Revisiting and Summarizing



### b. Visual Clarity



### c. Aesthetic Clarity



### d. Recommendations

